Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2019-04-05 Manuscript/manuscript.Rmd
manuscript edits

## MODELS IMPLEMENTED ON SMOTE RESAMPLED DATA PERFORMED BEST

Of the, resampling techniques considered, it was found that SMOTE resampled datasets, at the 10 kb resoltion, yielded the highest model performances across each metric. This result was, for the most part, consistent for the other data resolutions. Performances of randomly under-sampled datasets were found to be comparable to, if not better than (at some isolated resolutions; see supplementary), SMOTE. The steap drop off in performance for randomly over-sampled and no resampled datasets was most evident when assessing F1-scores and balanced accuracies. Measures of MCC were found to increase more steadily toward models implemented on SMOTE resampled datasets. These results imply that a best performing model, one that is able to best discriminate between genomic bins that are near versus far away from TAD boundaries, is one that incorporates both random under- and over-sampling, as SMOTE is known to do.

@@

@@

>>>>

@@

## VARIABLE IMPORTANCES FROM TOP PERFORMING MODELS CONFIRMED BIOLOGICAL FUNCTIONALITY OF KNOWN GENOMIC ANNOTATIONS

Lastly, we investigated which genomic annotations were most predictive of TAD boundary formation. Looking at heatmaps of variable importances provided by the random forest classifiers implemented on SMOTE re-sampled data, with distance-type predictors, we found that known transcription factors associated with the hypothesized loop extrusion model [??? include refs] were give high predictive importance in the model (Figure 10). These tended to include the SMC3ab, ZNF143, CTCF, and RAD21 transcription factor binding sites and were found to be ranked consistly across resolution. Other annotations known to be associated with gene regulation and affect structural properties of chromatin, such as the YY1 and MAZ transcription factors, as well DNAse hypersensitive sites, regions of heterochromatin, and the H3k27 trimethylation and H3K4me1 modifications were less consistently ranked across resolutions for either cell line.

@@

Although we were not able to draw direct comparisons between the GM12878 and K562 cell lines, given that they did not contain the same list of cell type specific annotations, we did find that there was relative agreement in rankings of the top 5-7 predictors. However, there were notable exclusions from each cell line, in addition to inconsistencies across resolutions
